Typestate Inference in an Imperative Calculus

### 2013-01-04 formalism/paper.tex
fixed compile error, minor layout problems in progress/preservation proof

\begin{thm}[Progress and Preservation]
\label{thm:soundness}
Given a term $t$ such that $\typerule{\Gamma}{t}{T}{\Gamma'}$
with a store $\mu$ such that $\Gamma \vdash \mu$, either $t$ is a value or 
there exists a $t'$ and $\mu'$ such that
$\oprule{t}{\mu}{t'}{\mu'}$, with some $\Gamma''$ such that
$\Gamma'' \vdash \mu'$ and $\typerule{\Gamma''}{t'}{T}{\Gamma'}$.
\end{thm}

\begin{proof}
by induction on the typing derivation of $t$. Assume $t$ is not a value. It has
therefore been typed by one of the following rules:

@@

  \begin{itemize}
  \item $t_v$ is a value.
  Reduction can occur by \rletv such that 
  $\oprule{\lett{x}{t_v}{t_b}}{\mu}{t_b}{\mu [ x \mapsto t_v ]}$.
  Let $\mu' = \mu [ x \mapsto t_v ]$. Let $\Gamma'' = \Gamma_1, x : T_v$.
  It follows that $\Gamma'' \vdash \mu'$ and by rule \tdrop that
  $\typerule{\Gamma''}{t_b}{T}{\Gamma'}$.

@@

  \begin{itemize}
  \item $t_l$ is a value. Reduction can therefore occur by
  \rseqlv such that
  $\oprule{t_l ; t_r}{\mu}{t_r}{\mu}$. 
  Let $\Gamma'' = \Gamma_{mid}$, and $\mu' = \mu$.
  Directly, $\typerule{\Gamma''}{t_r}{T}{\Gamma'}$.
  By \lemref{lem:ctxsubset} it follows that $dom(\Gamma'') \subseteq dom(\Gamma)$,
  and by \lemref{lem:valuectx} it follows that 
  $\forall y \in dom(\Gamma'') . \Gamma(y) <: \Gamma''(y)$.
  Therefore $\Gamma'' \vdash \mu'$.

>>>>

\begin{thm}[Progress and Preservation]
\label{thm:soundness}
Given a non-value term $t$ such that $\typerule{\Gamma}{t}{T}{\Gamma'}$
and a store $\mu$ such that $\Gamma \vdash \mu$, then:

\begin{itemize}
\item there exists a $t'$ and $\mu'$ such that
$\oprule{t}{\mu}{t'}{\mu'}$
\item there exists a $\Gamma''$ such that
$\Gamma'' \vdash \mu'$ and $\typerule{\Gamma''}{t'}{T}{\Gamma'}$
\end{itemize}
\end{thm}

\begin{proof}
by induction on the typing derivation of $t$. As $t$ is not a value,
$\typerule{\Gamma}{t}{T}{\Gamma'}$ must have been derived by one of the
following rules:

@@

  \begin{itemize}
  \item $t_v$ is a value.
  Reduction can occur by \rletv such that 
  $\oprule{t}{\mu}{t_b}{\mu [ x \mapsto t_v ]}$.
  Let $\mu' = \mu [ x \mapsto t_v ]$. Let $\Gamma'' = \Gamma_1, x : T_v$.
  It follows that $\Gamma'' \vdash \mu'$ and by rule \tdrop that
  $\typerule{\Gamma''}{t_b}{T}{\Gamma'}$.

@@

  \begin{itemize}
  \item $t_l$ is a value. Reduction can therefore occur by \\
  \rseqlv such that
  $\oprule{t_l ; t_r}{\mu}{t_r}{\mu}$. 
  Let $\Gamma'' = \Gamma_{mid}$, and $\mu' = \mu$.
  Directly, $\typerule{\Gamma''}{t_r}{T}{\Gamma'}$.
  By \lemref{lem:ctxsubset} it follows that $dom(\Gamma'') \subseteq dom(\Gamma)$,
  and by \lemref{lem:valuectx} it follows that 
  $\forall y \in dom(\Gamma'') . \Gamma(y) <: \Gamma''(y)$.
  Therefore $\Gamma'' \vdash \mu'$.
